---
title: "Antarctic_Rarefaction"
author: "Trevor Eakes"
date: "August 6, 2016"
output:
  html_document: default
  pdf_document: default
---
Rarefaction methods applied to Antarctic phytoplankton cell abundance counts taken from the raw data analized in Yamamoto, 1986. This analysis is faulty at the momement because it is to computationally challenging for the computer to manage, and the methods to not apply for how large the sample size is. See notes in script below.

###Note Diatoms in Antarc are tentatively from column 3 to 50

Preparing the data
```{r}
source('NifftyTrevFunctions.R')
Antarc <- read.csv("AntarcticOceandatasheet.csv") #Pull up datasheet
head(Antarc)
Antarc[,3] <-NULL #Get rid of diatoms column
Antarc[is.na(Antarc)] <- 0 #Get rid of pesky NA's. 
AntarcDiversity <- Create.Diversity.I(Antarc, 3, ncol(Antarc))#Use niffty function to 
#make diversity indices
View(AntarcDiversity)
SplitData(Antarc, "Location", "") #Splits up data into the 4 regions
#Antarctic
Antarctic <- Antarctic[, 3:ncol(Antarctic)] #take only the phytoplankton here 
Antarctic <- Antarctic[-which(Antarctic==0)] #get rid of 0's
Antarctic <- Antarctic[!is.na(Antarctic)] #get rid of NA's
#Middle
Middle <- Middle[, 3:ncol(Middle)]
Middle <- colSums(Middle) #add columns
Middle <- Middle[-which(Middle==0)]
Middle <- Middle[!is.na(Middle)]
#Subantarctic
Subantarctic <- Subantarctic[, 3:ncol(Subantarctic)]
Subantarctic <- colSums(Subantarctic)
Subantarctic <- Subantarctic[-which(Subantarctic==0)]
Subantarctic <- Subantarctic[!is.na(Subantarctic)]
#Subtropical
Subtropical <- Subtropical[, 3:ncol(Subtropical)]
Subtropical <- colSums(Subtropical)
Subtropical<- Subtropical[-which(Subtropical==0)]
Subtropical<- Subtropical[!is.na(Subtropical)]                                                    
```

### A heatmap plot of the abundance of different species in this data set.

```{r, fig.width=5, fig.height=25}
#Create big dataframe 
require(ggplot2)
require(gplots)
require(iNEXT) #Rarefaction package by Chao
#
Antarc2 <- Antarc[,3:ncol(Antarc)]
Antarc2<- rbind(Antarc2[1,], colSums(Antarc2[c(2,5,8),]), colSums(Antarc2[c(3,4,7),]), colSums(Antarc2[c(6,9),]))
Antarc2<- t(round(apply(Antarc2, 1, function(x) x/sum(x)), digits=4)) #simplified for easy plotting,
#now Antarc3 each abundance is a percentage of the total cells found per region
row.names(Antarc2) <- c("Antarctic", "Middle", "SubAntarctic", "Subtropical")
Antarc2 <- t(Antarc2)
Antarc2 <- Antarc2[order(Antarc2[,1]),decreasing=TRUE]
Antarc2[is.na(Antarc2)] <- 0
jet.colors <- colorRampPalette(c("#00007F", "#007FFF", "cyan", "#7FFF7F", "yellow",
"#FF7F00", "#7F0000"))
col <-jet.colors(48)
col <-c("white", col)
# (optional) defines the color breaks manually for a "skewed" color transition
col_breaks = c(seq(0, .001, length=1), 
               seq(.002, .01, length=10),
               seq(.011,.05,length=10),              
               seq(.051,.1,length=10),
               seq(.11, .25, length=9),
               seq(.251, .5, length=5),
               seq(.51, 1, length=5))
#Heatmap plotting function, just add matrix!
AbundHeatmap <- function(matrix, col, breaks) {
heatmap.2(matrix,
               cellnote = matrix, # same data set for cell labels
               offsetCol = 0.5,
               notecol="black",      # change font color of cell labels to black
               trace="none",      # turns off trace lines inside the heat map
               margins =c(9,9),# widens margins around plot
                cexCol=2.2,
                cexRow=1.3,
                srtCol=30,
               col=col,       # use on color palette defined earlier 
               breaks= breaks,    # enable color transition at specified limits
               dendrogram="none", # only draw a row dendrogram
               Colv=NA,              # turn off column clustering
               key=TRUE,
               key.title = NULL,
               symbreaks=TRUE,
               symkey=FALSE,
               density.info="none",
               densadj = 0.25,
               Rowv = FALSE,
               #denscol="black",
               keysize=.2,
               #( "bottom.margin", "left.margin", "top.margin", "left.margin" )
               #key.par=list(mar=c(7,.5,3,1)),
                #lmat -- added 2 lattice sections (5 and 6) for padding
               lmat=rbind( c(0, 3, 0), c(2,1,0), c(0,4,0) ), 
          lhei=c(.5, 5, .3),
          lwid = c(.5,3,.5))}
par(mar = rep(1, 4))
AbundHeatmap(Antarc2, col=col, breaks=col_breaks)


```
###Rarefaction of the prepared data

####Note: rarefaction seems problematic here

This script takes a very long time to run because of the high number of cells counted in the Antarctic samples. The rarefaction index may be working but likely does not give meaningful results at this scale. Asymptope of the rarefaction curve is reached very quickly relative to the total number of individuals. Consideration of the method, scaling, and aplicability should be made here. Try running the script overnight, if it is still not working a subsampling aproach of the population may need to be designed. 

##Script turned off for functionality

AntarcR <- list("Antarctic"=Antarctic, "Subantarctic"=Subantarctic, "Middle"=Middle, "Subtropical"=Subtropical)
#
#INEXT creates rarefaction dataframe and metadata for a group of samples
A0 <- iNEXT(AntarcR, q=0,  datatype="abundance") #richness 
A1 <- iNEXT(AntarcR, q=1,  datatype="abundance") #Shannon based diversity

#Rarefaction of abundance data showing richness
AntarcRareRichness <- ggiNEXT(A0, type = 1, facet.var = "none", color.var = "site") 
AntarcRareRichness+ylab("Richness")
#Rarefaction of abundance data showing diversity (H1)
AntarcRareH <- ggiNEXT(A1, type = 1, facet.var = "none", color.var = "site") 
AntarcRareH+ylab("1st order Hill #")
#Coverage plotted for 1st order hill number
AntarcRareCoverage <- ggiNEXT(A1, type = 2, facet.var = "none", color.var = "site")
AntarcRareCoverage+ylab("Sample coverage H1")
```
