---
title: "Clustering_Rareification_Heatmap"
author: "Trevor Eakes"
date: "May 8, 2016"
output:
  pdf_document: default
  html_document: default
---
##Rarefaction Methods for the Kuroshio

####Rarefaction is a method for comparing diversity between sample groups which may have different coverage, area, or even distributions. It estimates the relationship between how diversity such as richness or the 1st order Hill number accumulates as more individuals/samples are collected. Recent development of this package and the idea behind rarefaction now allow for using abundance in adition to frequency. Previously I could only use frequency with this idea which is statistically less powerful. 

Research on rarefaction conducted by:
Dr. Anne Chao
National Tsing Hua University
Ecological Statistics, biodiversity

##First to prepare the data
Note: You must first run the dataframe cunstruction Rmd chunks.
The file is titled DiversityDataframeConstruction.Rmd

You will also need the following packages downloaded: 
gplots
ggplots2
iNEXT

###Dataframe creation and visualization

####Note, other graphs may be created from the data easily by changing the inputs in this script. Only relavent graphs have been chosen for printing. 
```{r}
centitest <- read.csv('centitest.csv') #Pull up processed phytoplankton cell 
#counts csv from directory
Adiv.abiotic <- read.csv('Adiv.abiotic.csv')# Pull up all 
#other related environmental and diversity data
centitest.Clade <- cbind(clades=Adiv.abiotic$cluster, centitest)
#adding the clade t centitest

#Clustering the Fillament manually by salinity and position
Adiv.abioticR <- cbind("rows"=c(1:nrow(Adiv.abiotic)), Adiv.abiotic) 
# add rows to keep track of things
Fillament <- Adiv.abioticR[which(Adiv.abioticR[, "DfromF"]>0),]
#Take stations north of the front


#More manual editing, WILL NEED TO ADJUST THIS AS EASY,
#MAYBE SET UP OBJECTIVE FILTERING ALGORYTHM 
Fillament <- Fillament[which(Fillament[,"S"]>33.85),] #Take stations 
#Above 33.9 salinity
Fillament <- Fillament[which(Fillament["lat"]>35.99),]
Fillament <- Fillament[which(Fillament["station"]!=2),]
# must manually input single station number you want removed each time 
Fillament <- Fillament[which(Fillament["station"]!=3),] 
Fillament <- Fillament[which(Fillament["station"]!=2),]
Fillament <- Fillament[which(Fillament["depth..m."]<80),]

#vISUALIZE FILLAMENT STATIONS
require(ggplot2)
jet.colors <- colorRampPalette(c("#00007F", "#007FFF", "cyan", "#7FFF7F", "yellow",
"#FF7F00", "#7F0000"))

ggplot(Fillament, aes(x=lon, y=lat, colour=Theta))+ 
geom_point(size=10, position=position_dodge(), stat="identity")+
scale_color_gradientn(colours=jet.colors(7), space="rgb", guide="colourbar")
#
ggplot(Fillament, aes(x=lat, y=Theta, colour=S)) + geom_point(size=10,              position=position_dodge(), stat="identity")+
scale_color_gradientn(colours=jet.colors(7), space="rgb", guide="colourbar")
#VISUALIZE WHERE FILLAMENT IS RELATIVE TO ALL OTHER SAMPLES
Fill <- data.frame(rows=Fillament$rows, class=rep("Fill", length(Fillament$rows)))
Else <- Adiv.abioticR[-c(Fillament$rows), ]
Else <- data.frame(rows=Else$rows,  class=rep("Else", length(Else$rows)))
class <- rbind(Else, Fill)
class <- class[order(class$rows),]
Adiv.abioticR <- cbind(Adiv.abiotic, class)
#
ggplot(Adiv.abioticR, aes(x=lat, y=depth..m., colour=lon, shape=factor(class)))+
geom_point(size=3,position=position_dodge(), stat="identity")+
scale_color_gradientn(colours=jet.colors(7), space="rgb", guide="colourbar")

```

###Setting up the fillament data and preparing for rarefaction

```{r}

  SplitData <- function(data, column, header){
  #data=dataframe, column=the the number of
  #the column desired to be used as a numerical factor,
#header is the title of the data frames without the numbers at the end.
#MUST BE IN PARATHESES
libr <- setNames(split(data, data[, column]), paste0(header, unique(data[,column])))
#Seperates the dataframe into multiple dataframes based on the factor
  return(libr) #Exports both s and each individual, new data frame to the global environment 
  }
fillament <- centitest.Clade[Fillament$rows,]
#Grab data for only fillament stations
centitest.Clade <- centitest.Clade[-Fillament$rows,] 
#remove fillament stations from centitest.Clade
SplitD <- SplitData(centitest.Clade, 1, 'dfcluster') 
#Split up the data into the remaining clasified water masses
dfcluster1 <- SplitD$dfcluster1
dfcluster2 <-SplitD$dfcluster2
dfcluster3 <- SplitD$dfcluster3
#sums species in each cluster
Clus1 <- as.numeric(apply(dfcluster1[,6:ncol(dfcluster1)], 2, function(x) sum(x>0)))
#gets rid of the 0's, add the number of rows to the front to match the package format
clus1 <- c(nrow(dfcluster1), Clus1[Clus1>0])#And so on
Clus2 <- as.numeric(apply(dfcluster2[,6:ncol(dfcluster2)], 2, function(x) sum(x>0)))
clus2 <- c(nrow(dfcluster2), Clus2[Clus2>0])
Clus3 <- as.numeric(apply(dfcluster3[,6:ncol(dfcluster3)], 2, function(x) sum(x>0)))
clus3 <- c(nrow(dfcluster3), Clus3[Clus3>0])
Clus4 <- as.numeric(apply(fillament[,6:ncol(fillament)], 2, function(x) sum(x>0)))
clus4 <- c(nrow(fillament), Clus4[Clus4>0])

#For abundance, ABUNDANCE MOST INFORMATIVE
clusA1 <- as.numeric(apply(dfcluster1[,6:ncol(dfcluster1)], 2, function(x) sum(x))) 
#sum species abundances
clusA2 <- as.numeric(apply(dfcluster2[,6:ncol(dfcluster2)], 2, function(x) sum(x)))
clusA3 <- as.numeric(apply(dfcluster3[,6:ncol(dfcluster3)], 2, function(x) sum(x)))
clusA4 <- as.numeric(apply(fillament[,6:ncol(fillament)], 2, function(x) sum(x)))

#Making Rarefaction lists for the clusters with row names
Rarefaction <- list("Deep"=clus1, "Fillament"=clus4, "Kuroshio"=clus2, "Oyashio"=clus3) 
RarefactionAbund <- list("Deep"=clusA1, "Fillament"=clusA4, "Kuroshio"=clusA2, "Oyashio"=clusA3)
```

###List of stations and their depths

```{r}
#Summarizing Function, returns mean, standard deviation, range, and # of observations
summarize <- function(data) {
Means <- apply(data, 2, function(x) mean(x))
SD <- apply(data, 2, function(x) sd(x))
Min<-apply(data, 2, function(x) min(x))
Max<-apply(data, 2, function(x) max(x))
Obs<- apply(data, 2, function(x) length(x))
summry <- cbind(Means, SD, Min, Max, Obs)
rownames(summry) <- colnames(data)
return(summry)
}
#
WaterM <- cbind(clades=rep(4, length(Fillament$rows)), X=Fillament$rows)
#add code for fillament
WaterM <- rbind(WaterM, centitest.Clade[, 1:2]) #combine with other clade data
WaterM <- WaterM[order(WaterM[,2]),] #Reorder by row number
Adiv.abioticC <- cbind(WaterM, Adiv.abiotic) #add back to the dataframe

#Summary
CladeAdiv <- SplitData(Adiv.abioticC, "clades", "Adiv")
DeepAdiv <- CladeAdiv$Adiv1
FillAdiv <- CladeAdiv$Adiv2
OyaAdiv <- CladeAdiv$Adiv3
KuroAdiv <- CladeAdiv$Adiv4
(DeepSmry <- summarize(DeepAdiv[,c("Chlorophyll", "Theta", "S", "depth..m.", "Richness",
"Cellcount", "ShannonWiener", "sigPoDen")]))
(FillSmry <- summarize(FillAdiv[,c("Chlorophyll", "Theta", "S", "depth..m.", "Richness",
"Cellcount", "ShannonWiener", "sigPoDen")]))
(OyaSmry <- summarize(OyaAdiv[,c("Chlorophyll", "Theta", "S", "depth..m.", "Richness",
"Cellcount", "ShannonWiener", "sigPoDen")]))
(KuroSmry <- summarize(KuroAdiv[,c("Chlorophyll", "Theta", "S", "depth..m.", "Richness",
"Cellcount", "ShannonWiener", "sigPoDen")]))
#
#Number of Species which appear once or twice for each cluster
Kurolets <- length(which(Clus1<3 & Clus1>0)) #Kuroshio samples
Oyalets <- length(which(Clus2<3 & Clus2>0))  #Oyashio samples
Deeplets <- length(which(Clus3<3 & Clus3>0)) #Deep samples
Fillamentlets <- length(which(Clus4<3 & Clus4>0)) #Fillament
#
#Fillament Station info
(Fillamentinfo <- Adiv.abiotic[Fillament$rows, c("depth..m.", "station", "lat", "lon") ])
#Other water masses info
(KuroshioInfo <-Adiv.abiotic[dfcluster1$X,c("depth..m.", "station", "lat", "lon") ])
(OyashioInfo <- Adiv.abiotic[dfcluster2$X,c("depth..m.", "station", "lat", "lon") ])
(DeepInfo <-Adiv.abiotic[dfcluster3$X,c("depth..m.", "station", "lat", "lon") ])


```
##Heatmap ploting

Finally, We visualize how the communities appear in each clustered water mass, this can tell us why the water masses are considered different and in what way they difer in species abundance and presence

This is an important figure for understanding the Kuroshio 

Values are based off of phytoplankton incidence (# of times observed) which has been divided by the total number of samples in each group. We also use raw abundance data from each cluster to visualize actual cell counts. 


```{r fig.width = 5, fig.height = 12}
require(gplots)# for heatmap.2
library(RColorBrewer)# for better color options

#rounding count data
#For rounding frequency count data and standardizing it to the sample size
#Clus1, Clus2, Clus3, & Clus4 objects from the previous chunk contain the incidence
#information for each cluster.
#dfcluster# is the dataframe from which the centiliter abundance species sums 
#for the water mass were calculated and has the number of samples in rows
clus1c <- round(Clus1/nrow(dfcluster1[,-c(1:2)]), 2)*100
clus2c <- round(Clus2/nrow(dfcluster2[,-c(1:2)]), 2)*100
clus3c <- round(Clus3/nrow(dfcluster3[,-c(1:2)]), 2)*100
clus4c <- round(Clus4/nrow(fillament[,-c(1:2)]), 2)*100


Incidence <- as.matrix(data.frame(cbind(Kuroshio=clus1c, Filament=clus4c, 
Oyashio=clus3c, Deep=clus2c))) #combining into single data frame
Kuroshio_Phytoplankton <- read.csv("Kuroshio_Phytoplankton.csv")
#grab the original phytoplankton abundances dataframe
Kuroshio_Phytoplankton <- Kuroshio_Phytoplankton[-c(97:99, 186:190), -c(1,2)] 
#Editing base data frame, getting rid of uneaded columns and rows

#Raw Abundance Data
HeatAbund <- as.matrix(as.data.frame(RarefactionAbund))
#changing to a dataframe
row.names(HeatAbund)<-colnames(Kuroshio_Phytoplankton) 
#changing row names of Incidence to their true Phytoplankton names
row.names(Incidence)<-colnames(Kuroshio_Phytoplankton)
#
HeatAbund<-HeatAbund[order(HeatAbund[,3],decreasing=TRUE),]
#ordering rows from greatest to smallest
Incidence<-Incidence[order(Incidence[,3],decreasing=TRUE),]
#color scheme
col <-jet.colors(398)
col <-c("white", col)
# (optional) defines the color breaks manually for a "skewed" color transition
col_breaks = c(seq(0, 1, length=50), 
               seq(2, 3, length=50),
               seq(4,8,length=50),              
               seq(9,15,length=50),
               seq(16, 25, length=50),
               seq(26, 50, length=50),
               seq(51, 75, length=50),
               seq(76, 150, length=50))   
#Heatmap plotting function, just add matrix!
AbundHeatmap <- function(matrix, col, breaks) {
heatmap.2(matrix,
               cellnote = matrix, # same data set for cell labels
               offsetCol = 0.5,
               notecol="black",      # change font color of cell labels to black
               trace="none",      # turns off trace lines inside the heat map
               margins =c(8,8),     # widens margins around plot
               col=col,       # use on color palette defined earlier 
               breaks= breaks,    # enable color transition at specified limits
               dendrogram="none", # only draw a row dendrogram
               Colv=NA, # turn off column clustering
               cexCol=2.2,
               cexRow=1.3,
               srtCol=30,
               key=TRUE,
               key.title = NULL,
               symbreaks=TRUE,
               symkey=FALSE,
               density.info="none",
               densadj = 0.25,
               Rowv = FALSE,
               #denscol="black",
               keysize=.5,
               #( "bottom.margin", "left.margin", "top.margin", "left.margin" )
               #key.par=list(mar=c(7,.5,3,1)),
                #lmat -- added 2 lattice sections (5 and 6) for padding
               lmat=rbind( c(0, 3, 0), c(2,1,0), c(0,4,0) ), 
          lhei=c(.5, 5, .3),
          lwid = c(.5,3,.5))}
par(mar = rep(1, 4))
AbundHeatmap(HeatAbund, col=col, breaks=col_breaks) #Raw Abundance data
#new breaks for Incidence
Inc_col_breaks = c(seq(0, 1, length=50), seq(2, 3, length=50),seq(4,5,length=50),
seq(6,8,length=50),seq(9, 15,
length=50),seq(16, 20, length=50), seq(21, 30, length=50), seq(31, 45, length=50))   
AbundHeatmap(Incidence, col=col, breaks=Inc_col_breaks)
```

###Rarefaction and graphing Rarefaction curves

Note: 
Incidence rarefaction is turned on for improved speed running Rmd. Due to recent package updates incidence input data needs to be adjusted to match apropriate package format.

Aditional plotting may be done here using the INEXT package. This just shows the most relavent plots of interest 

```{r}
require(iNEXT) #Rarefaction package by Chao
#INEXT creates rarefaction dataframe and metadata for a group of samples
A0 <- iNEXT(RarefactionAbund, q=0,  datatype="abundance") #richness 
A1 <- iNEXT(RarefactionAbund, q=1,  datatype="abundance") #Shannon based diversity
#
#z0 <- iNEXT(Rarefaction, q=0,  datatype="Incidence_freq") #Incidence counts
#z1 <- iNEXT(Rarefaction, q=1,  datatype="incidence") #Incidence counts

#Rarefaction of abundance data showing richness
RarefactionRichness <- ggiNEXT(A0, type = 1, facet.var = "none", color.var = "site") 
RarefactionRichness+ylab("Richness")
#Rarefaction of abundance data showing diversity (H1)
RarefactionH <- ggiNEXT(A1, type = 1, facet.var = "none", color.var = "site") 
RarefactionH+ylab("1st order Hill #")
#Coverage plotted for 1st order hill number
RarefactionCoverage <- ggiNEXT(A1, type = 2, facet.var = "none", color.var = "site")
RarefactionCoverage+ylab("Sample coverage H1")
#Creates rarefaction graphs
```




